Community-conditioned language models for linguistic variation

### 2021-10-14 paper/paper1.tex
revision in sec 4

\subsection{Comparing embeddings: anecdotal evidence}
\label{sec:storytelling}

@@

Let us call \(L\) the matrix of (normed) linguistic embeddings and
\(S\) the matrix of (normed) social embeddings. Thus \(L_i\) and
\(S_i\) are the linguistic and social embeddings for community \(i\).

A straightforward, but ultimately flawed, way to measure 
how similar the two spaces are is to
consider each pair of communities \((i,j)\), and compute the
correlation between the cosine similarities of both embeddings. That
is, compute the Pearson correlation factor of the data set:
\[C = \{(x = L_i · L_j, y = S_i · S_j) \text {~for~} i,j ∈ [1,510]\}\] % BN: changed \C to C. Definition for \C may be missing

Indeed, $C$ shows positive correlation for both the LSTM ($r=0.438$)
and transformer ($r=0.452$) linguistic embeddings.\footnote{
  By comparison, the correlation between the two linguistic embeddings is \num{0.759}.}
The correlations are significant with $p<0.001$ in all cases;
however, we note that the number of pairs grows with the square
of the number of communities (with 510 communities, we have 129795)
pairs), meaning that standard statistical tests on Pearson correlation will
assure us of statistical significance in all
but the weakest of correlations. A further flaw is that the data points in 
\(C\) are \emph{not} distributed independently---far from it since each data point is generated 
from \num{2} of \num{510} independent variables. We consider this last flaw fatal, 
and take a different approach for computing the correlation between community
embeddings (\cref{sec:statistical-evidence}).

We can, however, use cosine similarity to find pairs of communities
that are similar in one embedding but different in the other. 
This gives us a way of characterizing the differences between the two kinds of
community representation.
The examples shown in \cref{fig:embedding-differences-examples} were selected by taking the \num{5} 
most linguistically similar (in the LSTM embedding) pairs of communities that had below median social similarity 
(and \textit{vice versa}).\footnote{
Median similarity among pairs of communities was \num{0.177} for the social embedding and
\num{0.010} and \num{0.012} for the LSTM and transformer linguistic embeddings, respectively.}

@@

Differences at the other end of the spectrum (\cref{fig:embedding-differences-examples}, right)
are somewhat harder to interpret.
It is mostly easy to see why these communities would have different linguistic embeddings---%
in all cases the topics are quite different.
The reason they have similar social embeddings is less obvious,
but we can discern some trends in how the communities are premised.
The \subreddit{progresspics} and \subreddit{TalesFromRetail} are premised, in part,
on seeking support from other people with similar experiences;
\subreddit{legaladvice}, \subreddit{Cooking}, and \subreddit{loseit}
all have a knowledge-sharing component to them;
\subreddit{running} and \subreddit{Coffee} are hobby-focused;
and \subreddit{self} (often) and \subreddit{askscince} (by premise) are
places people ask and answer questions.
It may be that there are different patterns in the \emph{social function}
that people attribute to this particular social media website---%
people who use Reddit in one way are more likely
to belong to communities that are premised on the same kind of social function,
even if the topics (and indeed language) of those communities are quite different.
Testing this hypothesis would require a more focused study design and 
ideally consider communities from multiple social networks (online or otherwise).

\subsection{Comparing embeddings: statistical evidence}
\label{sec:statistical-evidence}

>>>>

\subsection{Comparing embeddings: cosine similarities}
\label{sec:storytelling}

@@

A difficulty is that the social embeddings and linguistic embeddings
range over completely unrelated spaces, so one cannot use the usual
cosine similarity metric \emph{across} this boundary. One can,
however, use cosine similarity between \emph{pairs} of communities,
and verify the similarities correlated between linguistic to social
embeddings.  This gives us a way of characterizing the differences
between the two kinds of community representation. To get a more
concrete sense of what this method yields, we first survey some of the
most salient community pairs.

\todo{insert some examples with very high correlation}

The examples shown in \cref{fig:embedding-differences-examples} were
selected by taking the \num{5} most linguistically similar (in the
LSTM embedding) pairs of communities that had below median social
similarity (and \textit{vice versa}).\footnote{ Median similarity
  among pairs of communities was \num{0.177} for the social embedding
  and \num{0.010} and \num{0.012} for the LSTM and transformer
  linguistic embeddings, respectively.}

@@

Differences at the other end of the spectrum
(\cref{fig:embedding-differences-examples}, right) are somewhat harder
to interpret.  It is mostly easy to see why these communities would
have different linguistic embeddings---%
in all cases the topics are quite different.  The reason they have
similar social embeddings is less obvious, but we can discern some
trends in how the communities are premised.  The
\subreddit{progresspics} and \subreddit{TalesFromRetail} are premised,
in part, on seeking support from other people with similar
experiences; \subreddit{legaladvice}, \subreddit{Cooking}, and
\subreddit{loseit} all have a knowledge-sharing component to them;
\subreddit{running} and \subreddit{Coffee} are hobby-focused; and
\subreddit{self} (often) and \subreddit{askscience} (by premise) are
places people ask and answer questions.  It may be that there are
different patterns in the \emph{social function} that people attribute
to this particular social media website---%
people who use Reddit in one way are more likely to belong to
communities that are premised on the same kind of social function,
even if the topics (and indeed language) of those communities are
quite different.  Testing this hypothesis would require a more focused
study design and ideally consider communities from multiple social
networks (online or otherwise).

We simultaneously have examples of high correlation between social and
linguistic embeddings, but also of low correlation. This is
essentially anecdotal evidence: we must resort to statistical tools to
extract the general trends.
A straightforward, but ultimately flawed,
way to measure how similar the two spaces are is to consider each pair
of communities \((i,j)\), and compute the correlation between the
cosine similarities of both embeddings.

That is, we can compute the Pearson correlation factor of the data set:
\[C = \{(x = L_i · L_j, y = S_i · S_j) \text {~for~} i,j ∈
  [1,510]\}\] % BN: changed \C to C. Definition for \C may be missing
where \(L_i\) and \(S_i\) are the linguistic and social embeddings for
community \(i\).  (Thus \(L\) is the matrix of (normed) linguistic
embeddings and \(S\) the matrix of (normed) social embeddings.)

The analysis shows positive correlation for both the LSTM ($r=0.438$)
and transformer ($r=0.452$) linguistic embeddings.\footnote{ By
  comparison, the correlation between the two linguistic embeddings is
  \num{0.759}.}  The correlations are significant with $p<0.001$ in
all cases.  However, we note that the number of pairs grows with the
square of the number of communities (with 510 communities, we have
129795) pairs), meaning that standard statistical tests on Pearson
correlation will assure us of statistical significance in all but the
weakest of correlations. A further flaw is that the data points in
\(C\) are \emph{not} distributed independently---far from it in fact,
since each data point is generated from \num{2} of \num{510}
independent variables. We consider this last flaw fatal, and take a
different approach for computing the correlation between community
embeddings in the next section.

\subsection{Comparing embeddings: procrustes method}
\label{sec:statistical-evidence}
